---
title: "Cleaning and Analysis Set-Up"
author: "Cooperatives Working Group"
date: "last compilation: `r Sys.Date()`"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

# Load packages

```{r}
suppressPackageStartupMessages({
  library(here)
  library(raster)
  library(tmap)
  library(sf)
  library(scales)
  library(cowplot)
  library(tidyverse)
})
```


# Read the data in

For now, we are not including gear type and range of fishing activity because they make no sense.

## Read coordinate data

```{r}
coords <- read.csv(here("raw_data", "cooperative_coordinates.csv"),
                 stringsAsFactors = F,
                 strip.white = T) %>% 
  filter(!duplicated(.))
```


## Read number of species data

```{r}
spp_data <- read.csv(here("raw_data", "spp_by_cooperative.csv"),
                 stringsAsFactors = F,
                 strip.white = T)
```

## Load fishery type

```{r}
fish_type <- read.csv(here("raw_data", "fishtype.csv"),
                      stringsAsFactors = F,
                      strip.white = T)
```

## Read species scientific names

## Read species suceptibility (Jones & Cheung)
 

## Extract temperatures

```{r}
proj2 <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
data(World)
World <- as(World, "sf") %>% 
  sf::st_transform(proj2) %>% 
  mutate(N = 1) %>% 
  sf::st_union(by = N) %>% 
  sf::as_Spatial()
```


```{r}
# The script in scripts/change_in_temperature produces this RDS file. Run from command line for faster performance

r <- readRDS(file = here("data", "tsdiff.rds"))
```

```{r}
saltwater <- coords %>%
  left_join(fish_type, by = "fishery_id") %>% 
  filter(!fish_type %in% c("freshwater", "freshwater & diadromous")) %>%
  group_by(fishery_id) %>% 
  summarize(lon = mean(lon, na.rm = T), lat = mean(lat, na.rm = T)) %>% 
  rbind(data.frame(fishery_id = 0, lon = -31.427525, lat = -71.6063907)) %>% 
  filter(!is.na(lon))

xy <- data.frame(X = saltwater$lon, Y = saltwater$lat)
coordinates(xy) <- c("X", "Y")
proj4string(xy) <- proj2  ## for example
xy <- SpatialPointsDataFrame(xy, saltwater, proj4string = proj2)
```

```{r}
temps <- raster::extract(r, xy, buffer = 50000, fun = mean)
```

```{r}
coords <- cbind(xy, temps)
colnames(coords@data) <- c("fishery_id", "lon", "lat", "temperature_change")
coords <- coords@data
```

## Get all together


```{r}
coop <- read.csv(here("raw_data", "CoopResilience.csv"),
                 stringsAsFactors = F,
                 strip.white = T,
                 na.strings = c("N/A", "NA")) %>% 
  mutate(short_catch_use_numeric = case_when(short_catch_use == "subsistence" ~ 1,
                                             short_catch_use == "local market/subsistence" ~ 1,
                                             short_catch_use == "local market" ~ 1,
                                             TRUE ~ 0),
         short_catch_use_numeric = ifelse(is.na(short_catch_use), NA, short_catch_use_numeric),
         itq = str_detect(string = catch_share_types, pattern = "itq"),
         ivq = str_detect(string = catch_share_types, pattern = "ivq"),
         ieq = str_detect(string = catch_share_types, pattern = "iteq"),
         iq = str_detect(string = catch_share_types, pattern = "iq"),
         turf = str_detect(string = catch_share_types, pattern = "turf"),
         coop = str_detect(string = catch_share_types, pattern = "coop")) %>% 
    left_join(coords, by = "fishery_id") %>% 
  left_join(spp_data, by = c("original_order")) %>%
  left_join(fish_type, by = "original_order") %>% 
  filter(!fish_type %in% c("freshwater", "freshwater & diadromous")) %>% 
  dplyr::select(original_order, fishery_id, host_country, target_species, lon, lat, human_development_index, vessels_coop, processed, codified_penalties, cooperative_administration, direct_enforcement, financial_services, information_sharing, percent_of_gdp_from_fishing, short_participants, short_catch_use_numeric, sea_level_vulnerability, sea_temp_vulnerability, ocean_acidification_vulnerability, ozone_uv_vulnerability, recovery_time, temporal_no_take, spatial_no_take, msc_certification, recorded_fishery_closure, poverty_index, institutional_stability, corruption, itq, ivq, ieq, iq, turf, coop, mpa, profit_sharing, coharvest, cooperative_gear_restrictions, cooperative_size_limit, gear_sharing, habitat_restoration, restocking, research_support, recovery_potential, temperature_change, number_of_species) %>% 
  mutate(sea_level_vulnerability = -sea_level_vulnerability,
         ozone_uv_vulnerability = -ozone_uv_vulnerability,
         sea_temp_vulnerability = -sea_level_vulnerability,
         ocean_acidification_vulnerability = - ocean_acidification_vulnerability,
         recovery_time = - recovery_time,
         corruption = -corruption,
         percent_of_gdp_from_fishing = -percent_of_gdp_from_fishing,
         poverty_index = -poverty_index,
         short_participants = -short_participants,
         temperature_change = -temperature_change)

coop$temperature_change[62] <- coords$temperature_change[coords$fishery_id == 0]

coop_text <- coop %>%
  dplyr::select(original_order, fishery_id, host_country, target_species, lon, lat)

coop_numbers <- coop %>%
  dplyr::select(-c(original_order, fishery_id, host_country, target_species, lon, lat)) %>% 
  mutate_all(as.numeric) %>% 
  mutate_all(rescale) %>% 
  mutate(add_catch_shares = ((itq + ivq + ieq + iq+ turf+coop) > 0) *1) 

```

```{r}
coop_clean <- cbind(coop_text, coop_numbers) %>%
  mutate(occupational_multiplicity = human_development_index,
         target_diversity = number_of_species,
         material_style_of_life = processed, 
         social_capital = (add_catch_shares +
                             mpa +
                             temporal_no_take +
                             spatial_no_take +
                             msc_certification + 
                             institutional_stability + 
                             cooperative_administration + 
                             direct_enforcement
                             ) / 7,
         subsidies = financial_services,
         trust = information_sharing,
         economic_dependence = percent_of_gdp_from_fishing,
         subsistance_dependence = short_catch_use_numeric,
         habitat_susceptibility = (sea_level_vulnerability + 
                                     sea_temp_vulnerability + 
                                     ocean_acidification_vulnerability +
                                     ozone_uv_vulnerability) / 4,
         overexploitation = recorded_fishery_closure) %>% 
  mutate(ecological_exposure = temperature_change,
         ecological_recovery_potential = recovery_potential,
         social_adaptive_capacity = (occupational_multiplicity + 
                                       target_diversity + 
                                       social_capital +
                                       subsidies + 
                                       trust +
                                       material_style_of_life
                                     ) / 5,
         social_sensitivity = (economic_dependence +
                                 subsistance_dependence) / 2,
         ecological_sensitivity = (habitat_susceptibility + 
                                     overexploitation) / 2, # This is missing spp suscep
         ecological_vulnerability = ecological_exposure + ecological_sensitivity - ecological_recovery_potential) %>% 
  mutate(score = ecological_vulnerability + social_sensitivity - social_adaptive_capacity)

coop_clean %>% dplyr::select(score) %>% drop_na() %>% dim()

coop_clean %>% filter(!is.na(score)) %>% dplyr::select(fishery_id) %$% unique(fishery_id) %>% length()
```

```{r, fig.height = 6, fig.width = 8}
World2 <- sf::st_as_sf(World)

coop_clean %>% 
  dplyr::select(lon, lat, social_adaptive_capacity, social_sensitivity, ecological_sensitivity, ecological_exposure, ecological_recovery_potential) %>% 
  gather(variable, value, -c(lon, lat)) %>%
  drop_na() %>% 
  ggplot() +
  geom_sf(data = World2) +
  geom_point(aes(x = lon, y = lat, fill = value), shape = 21, size = 2) +
  theme_bw() +
  scale_fill_gradientn(colours = colorRamps::matlab.like(20)) +
  facet_wrap(~variable, ncol = 2)

coop_clean %>% 
  dplyr::select(lon, lat, score) %>% 
  drop_na() %>% 
  ggplot() +
  geom_sf(data = World2) +
  geom_point(aes(x = lon, y = lat, fill = score), shape = 21, size = 2) +
  theme_bw() +
  scale_fill_gradientn(colours = colorRamps::matlab.like(20))
```

```{r}
library(corrplot)

coop_clean %>% 
  dplyr::select(social_adaptive_capacity, social_sensitivity, ecological_sensitivity, ecological_exposure, ecological_recovery_potential) %>% 
  drop_na() %>% 
  as.matrix() %>% 
  cor() %>% 
  corrplot(type = "lower", method = "ellipse", diag = F)
```

```{r}
pca_data <- coop_clean %>% 
  magrittr::set_rownames(value = paste(.$original_order, .$fishery_id, .$host_country, sep = "-")) %>% 
  dplyr::select(human_development_index,
                number_of_species,
                processed,
                add_catch_shares,
                mpa,
                temporal_no_take,
                spatial_no_take,
                msc_certification,
                institutional_stability,
                cooperative_administration,
                direct_enforcement,
                financial_services,
                information_sharing,
                percent_of_gdp_from_fishing,
                short_catch_use_numeric,
                sea_level_vulnerability,
                sea_temp_vulnerability,
                ocean_acidification_vulnerability,
                ozone_uv_vulnerability,
                recorded_fishery_closure,
                ecological_exposure,
                ecological_recovery_potential)

pca_data %>% 
  drop_na() %>% 
  cor() %>% 
  corrplot::corrplot(type = "lower", method = "ellipse", diag = F)
```

## PCA for Social

```{r}
pca_data %>% 
  dplyr::select(human_development_index,
                number_of_species,
                processed,
                add_catch_shares,
                mpa,
                temporal_no_take,
                spatial_no_take,
                msc_certification,
                institutional_stability,
                cooperative_administration,
                direct_enforcement,
                financial_services,
                information_sharing,
                percent_of_gdp_from_fishing,
                short_catch_use_numeric) %>% 
  drop_na() %>% 
  as.matrix() %>% 
  prcomp() %>% 
  ggbiplot::ggbiplot(obs.scale = 1, var.scale = 1, circle = TRUE)
```

```{r}
pca_data %>% 
  dplyr::select(sea_level_vulnerability,
                sea_temp_vulnerability,
                ocean_acidification_vulnerability,
                ozone_uv_vulnerability,
                recorded_fishery_closure,
                ecological_exposure,
                ecological_recovery_potential) %>% 
  drop_na() %>% 
  as.matrix() %>% 
  prcomp() %>% 
  ggbiplot::ggbiplot(obs.scale = 1, var.scale = 1, circle = TRUE)
```

# How many points would we have by removing a single variable?

```{r}
nas <- numeric(length = dim(pca_data)[2])

for (i in 1:length(nas)){
  testing <- pca_data[,-i]
  nas[i] <- dim(pca_data)[1] - sum(apply(testing, 1, function(x){any(is.na(x))}))
}

nas %>% 
  magrittr::set_names(value = colnames(pca_data)) %>% 
  as.data.frame() %>% 
  magrittr::set_colnames(value = "points")
```


```{r}
coop_clean <- coop_clean %>% 
  dplyr::select(-c(7:48))

write.csv(coop_clean, file = here::here("data", "clean_cooperatives_data.csv"), row.names = F)
```


##### Option B

```{r}
coop_clean <- cbind(coop_text, coop_numbers) %>%
  mutate(occupational_multiplicity = human_development_index,
         target_diversity = number_of_species,
         material_style_of_life = processed, 
         social_capital = (#add_catch_shares +
                             # mpa +
                             # temporal_no_take +
                             # spatial_no_take +
                             msc_certification + 
                             institutional_stability + 
                             cooperative_administration # + 
                             # direct_enforcement
                             ) / 3,
         subsidies = financial_services,
         trust = information_sharing,
         economic_dependence = percent_of_gdp_from_fishing,
         subsistance_dependence = short_catch_use_numeric,
         habitat_susceptibility = (sea_level_vulnerability + 
                                     sea_temp_vulnerability + 
                                     ocean_acidification_vulnerability +
                                     ozone_uv_vulnerability) / 4,
         overexploitation = recorded_fishery_closure) %>% 
  mutate(ecological_exposure = temperature_change,
         ecological_recovery_potential = recovery_potential,
         social_adaptive_capacity = (occupational_multiplicity + 
                                       target_diversity + 
                                       social_capital +
                                       subsidies #+
                                       # trust +
                                       # material_style_of_life
                                     ) / 4,
         social_sensitivity = (economic_dependence +
                                 subsistance_dependence) / 2,
         ecological_sensitivity = (habitat_susceptibility + 
                                     overexploitation) / 2, # This is missing spp suscep
         ecological_vulnerability = ecological_exposure + ecological_sensitivity - ecological_recovery_potential) %>% 
  mutate(score = ecological_vulnerability + social_sensitivity - social_adaptive_capacity)

coop_clean %>% dplyr::select(score) %>% drop_na() %>% dim()

coop_clean %>% filter(!is.na(score)) %>% dplyr::select(fishery_id) %$% unique(fishery_id) %>% length()
```

```{r, fig.height = 6, fig.width = 8}
World2 <- sf::st_as_sf(World)

coop_clean %>% 
  dplyr::select(lon, lat, social_adaptive_capacity, social_sensitivity, ecological_sensitivity, ecological_exposure, ecological_recovery_potential) %>% 
  gather(variable, value, -c(lon, lat)) %>%
  drop_na() %>% 
  ggplot() +
  geom_sf(data = World2) +
  geom_point(aes(x = lon, y = lat, fill = value), shape = 21, size = 2) +
  theme_bw() +
  scale_fill_gradientn(colours = colorRamps::matlab.like(20)) +
  facet_wrap(~variable, ncol = 2)

coop_clean %>% 
  dplyr::select(lon, lat, score) %>% 
  drop_na() %>% 
  ggplot() +
  geom_sf(data = World2) +
  geom_point(aes(x = lon, y = lat, fill = score), shape = 21, size = 2) +
  theme_bw() +
  scale_fill_gradientn(colours = colorRamps::matlab.like(20))
```

```{r}
coop_clean %>% 
  dplyr::select(social_adaptive_capacity, social_sensitivity, ecological_sensitivity, ecological_exposure, ecological_recovery_potential) %>% 
  drop_na() %>% 
  as.matrix() %>% 
  cor() %>% 
  corrplot(type = "lower", method = "ellipse", diag = F)
```

```{r}
pca_data <- coop_clean %>% 
  magrittr::set_rownames(value = paste(.$original_order, .$fishery_id, .$host_country, sep = "-")) %>% 
  dplyr::select(human_development_index,
                number_of_species,
                # processed,
                # add_catch_shares,
                # mpa,
                # temporal_no_take,
                # spatial_no_take,
                msc_certification,
                institutional_stability,
                cooperative_administration,
                # direct_enforcement,
                financial_services,
                # information_sharing,
                percent_of_gdp_from_fishing,
                short_catch_use_numeric,
                sea_level_vulnerability,
                sea_temp_vulnerability,
                ocean_acidification_vulnerability,
                ozone_uv_vulnerability,
                recorded_fishery_closure,
                ecological_exposure,
                ecological_recovery_potential)

pca_data %>% 
  drop_na() %>% 
  cor() %>% 
  corrplot::corrplot(type = "lower", method = "ellipse", diag = F)
```

## PCA for Social

```{r}
pca_data %>% 
  dplyr::select(human_development_index,
                number_of_species,
                # processed,
                # add_catch_shares,
                # mpa,
                # temporal_no_take,
                # spatial_no_take,
                msc_certification,
                institutional_stability,
                cooperative_administration,
                # direct_enforcement,
                financial_services,
                # information_sharing,
                percent_of_gdp_from_fishing,
                short_catch_use_numeric) %>% 
  drop_na() %>% 
  as.matrix() %>% 
  prcomp() %>% 
  ggbiplot::ggbiplot(obs.scale = 1, var.scale = 1, circle = TRUE)
```

```{r}
pca_data %>% 
  dplyr::select(sea_level_vulnerability,
                sea_temp_vulnerability,
                ocean_acidification_vulnerability,
                ozone_uv_vulnerability,
                recorded_fishery_closure,
                ecological_exposure,
                ecological_recovery_potential) %>% 
  drop_na() %>% 
  as.matrix() %>% 
  prcomp() %>% 
  ggbiplot::ggbiplot(obs.scale = 1, var.scale = 1, circle = TRUE)
```

```{r}
pca_data %>% 
  drop_na() %>% 
  as.matrix() %>% 
  prcomp() %>% 
  ggbiplot::ggbiplot(obs.scale = 1, var.scale = 1, circle = TRUE)
```
















